---
title: "Lab3A"
author: "Alice Zheng"
date: "9/7/2020"
output: html_document
---
# Starting with Data
```{r setup}
library(tidyverse)
download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "data/portal_data_joined.csv")
```

```{r}
surveys <- read_csv("data/portal_data_joined.csv")
head(surveys)
```
```{r}
str(surveys)
```
## Challenge
* The object surveys contain the classes ‘spec_tbl_df’, ‘tbl_df’, ‘tbl’ and 'data.frame'.
* There are 34786 rows and 17 columns
* There are 48 species

```{r}
surveys[1,1]
surveys[1,6]
head_surveys <- surveys[1:6, ]
```
```{r, include=FALSE}
surveys$species_id
```

## Challenge
### 1.
```{r}
surveys_200 <- surveys[200, ]
surveys_200
```
### 2.
```{r}
nrow(surveys)
tail(surveys)
```
```{r}
surveys_last <- surveys[nrow(surveys), ]
surveys_last
```
### 3.
```{r}
surveys_middle <- surveys[nrow(surveys)/2, ]
surveys_middle
```
### 4.
```{r}
surveys[-(7:nrow(surveys)),]
```

## Factors
```{r}
sex <- factor(c("male", "female", "female", "male"))
levels(sex)
nlevels(sex)
sex
```

```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(levels(year_fct))[year_fct]
```

## Renaming Factors
```{r}
plot(as.factor(surveys$sex))
```
```{r}
sex <- factor(surveys$sex)
head(sex)
```

```{r}
levels(sex)
levels(sex)[0] <- "undetermined"
levels(sex)
head(sex)
```

## Challenge
```{r}
levels(sex)[1] <- "female"
levels(sex)[2] <- "male"
levels(sex)
head(sex)
```
```{r}
plot(sex)
```
 
## Using #StringAsFactor
```{r}
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = TRUE)
str(surveys)
```
```{r}
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = FALSE)
str(surveys)
```
```{r}
surveys$plot_type <- factor(surveys$plot_type)
```

## Challenge
```{r}
animal_data <- data.frame(
          animal = c("dog", "cat", "sea cucumber", "sea urchin"),
          feel = c("furry", "squishy", "spiny", "soft"),
          weight = c(45, 8, 1.1, 0.8)
          )
```
Needed quotes around the animals, a fourth feel, and a comma in the weight

```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, "15"),
       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1)
       )
str(country_climate)
```
Did not expect temperature and hemisphere to be factors, and has kangaroo to be numeric.

```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, "15"),
       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1)
       )
```

Using stringsAsFactors = FALSE would have made them characters instead of factors. Remove the quotes around temperature 15 and FALSE in hemisphere, and change the 1 in kangaroo to FALSE.

## Formatting Dates
```{r}
str(surveys)
library("lubridate")
```

```{r}
my_date <- ymd("2015-01-01")
str(my_date)
```
```{r}
my_date <- ymd(paste("2015", "1", "1", sep = "-")) 
str(my_date)
```

```{r, include=FALSE}
paste(surveys$year, surveys$month, surveys$day, sep = "-")
```
```{r, include=FALSE}
ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
```
```{r}
surveys$date <- ymd(paste(surveys$year, surveys$month, surveys$day, sep = "-"))
```
```{r}
str(surveys)
```
```{r}
summary(surveys$date)
```
```{r}
missing_dates <- surveys[is.na(surveys$date), c("year", "month", "day")]

head(missing_dates)
```
These dates failed to parse because they don't exist, and there are only 30 days in those months.
