---
title: "Lab4a"
author: "Alice Zheng"
date: "9/14/2020"
output: html_document
---
```{r}
library(tidyverse)
```

## Plotting with ggplot2
```{r}
surveys_complete <- read_csv("data/portal_data_joined.csv")
```
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
  geom_point()
```
```{r}
surveys_plot <- ggplot(data = surveys_complete, 
                       mapping = aes(x = weight, y = hindfoot_length))
```
```{r}
surveys_plot+
  geom_point()
```


## Challenge
```{r}
library("hexbin")
```
```{r}
surveys_plot +
 geom_hex()
```
A hexagonal is more useful when there is a high density of points in an area compared to a scatter plot. On the other hand if there are less points, a scatterplot is more precise. 

```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1)
```
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, color = "blue")
```
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, aes(color = species_id))
```
```{r}
str(surveys_complete)
```


## Challenge
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, aes(color = plot_type))
```
This is not a good way to show this type of data because the different colors are too mixed together and small to contribute any information to the graph. 

## Boxplot
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot()
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3, color = "tomato")
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_jitter(alpha = 0.3, color = "tomato")+
  geom_boxplot(alpha = 0)
```

## Challenge
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_jitter(alpha = 0.3, color = "tomato")+
  geom_violin()
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_jitter(alpha = 0.3, color = "tomato")+
  geom_violin()+
  scale_y_log10()
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, aes(color = plot_id)) +
  geom_boxplot(alpha = 0)
```

