---
title: "Lab4a"
author: "Alice Zheng"
date: "9/14/2020"
output: html_document
---
# Data Visualization
```{r}
library(tidyverse)
```

## Plotting with ggplot2
```{r}
surveys_complete <- read_csv("data/portal_data_joined.csv")
```
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
  geom_point()
```
```{r}
surveys_plot <- ggplot(data = surveys_complete, 
                       mapping = aes(x = weight, y = hindfoot_length))
```
```{r}
surveys_plot+
  geom_point()
```


## Challenge
```{r}
library("hexbin")
```
```{r}
surveys_plot +
 geom_hex()
```
A hexagonal is more useful when there is a high density of points in an area compared to a scatter plot. On the other hand if there are less points, a scatterplot is more precise. 

```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1)
```
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, color = "blue")
```
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, aes(color = species_id))
```
```{r}
str(surveys_complete)
```


## Challenge
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, aes(color = plot_type))
```
This is not a good way to show this type of data because the different colors are too mixed together and small to contribute any information to the graph. 

## Boxplot
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot()
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3, color = "tomato")
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_jitter(alpha = 0.3, color = "tomato")+
  geom_boxplot(alpha = 0)
```

## Challenge
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_jitter(alpha = 0.3, color = "tomato")+
  geom_violin()
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
  geom_jitter(alpha = 0.3, color = "tomato")+
  geom_violin()+
  scale_y_log10()
```
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) +
  geom_jitter(alpha = 0.3, aes(color = plot_id)) +
  geom_boxplot(alpha = 0)
```

## Plotting time series data
```{r}
library(tidyverse)
surveys_complete <- read_csv("data/portal_data_joined.csv")
```

```{r}
yearly_counts <- surveys_complete %>%
  count(year, genus)
ggplot(data = yearly_counts, aes(x = year, y = n)) +
     geom_line()
```
```{r}
ggplot(data = yearly_counts, aes(x = year, y = n, group = genus)) +
    geom_line()
```
```{r}
ggplot(data = yearly_counts, aes(x = year, y = n, color = genus)) +
    geom_line()
```

## Integrating the pipe operator with ggplot2
```{r}
yearly_counts %>% 
    ggplot(mapping = aes(x = year, y = n, color = genus)) +
    geom_line()
```
```{r}
yearly_counts_graph <- surveys_complete %>%
    count(year, genus) %>% 
    ggplot(mapping = aes(x = year, y = n, color = genus)) +
    geom_line()

yearly_counts_graph
```

## Faceting
```{r, fig.width=10,fig.height=11}
ggplot(data = yearly_counts, aes(x = year, y = n)) +
    geom_line() +
    facet_wrap(facets = vars(genus), ncol = 3)
```
```{r, fig.width=10,fig.height=11}
yearly_sex_counts <- surveys_complete %>%
                      count(year, genus, sex)
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_wrap(facets =  vars(genus), ncol=3)
```
```{r, fig.width=13,fig.height=7}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(sex), cols =  vars(genus))
```
```{r, fig.width=10,fig.height=15}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(genus))

```
```{r, fig.width=13,fig.height=6}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(cols = vars(genus))

```

## ggplot2 themes
```{r, fig.width=10,fig.height=11}
ggplot(data = yearly_sex_counts, 
        mapping = aes(x = year, y = n, color = sex)) +
     geom_line() +
     facet_wrap(vars(genus), ncol=3) +
     theme_bw()
```

## Challenge
```{r, fig.width=10,fig.height=14}
yearly_weight <- surveys_complete %>%
                group_by(year, species_id) %>%
                 summarize(avg_weight = mean(weight))
ggplot(data = yearly_weight, aes(x = year, y = avg_weight)) +
    geom_line() +
    facet_wrap(facets = vars(species_id), ncol=3)
```

## Customization
```{r, fig.width=10,fig.height=11}
ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus), ncol=3) +
    labs(title = "Observed genera through time",
         x = "Year of observation",
         y = "Number of individuals") +
    theme_bw()
```
```{r, fig.width=10,fig.height=11}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus), ncol=3) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(text=element_text(size = 16))
```

```{r, fig.width=10,fig.height=11}
ggplot(data = yearly_sex_counts, mapping = aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus), ncol=3) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 12),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16))
```

```{r, fig.width=10,fig.height=6}
grey_theme <- theme(axis.text.x = element_text(colour="grey20", size = 12, 
                                               angle = 90, hjust = 0.5, 
                                               vjust = 0.5),
                    axis.text.y = element_text(colour = "grey20", size = 12),
                    text=element_text(size = 16))

ggplot(surveys_complete, aes(x = species_id, y = hindfoot_length)) +
    geom_boxplot() +
    grey_theme

```

## Challenge
```{r, fig.width=10,fig.height=11}
yearly_counts %>% 
  ggplot(mapping = aes(x = year, y = n)) +
  facet_wrap(vars(genus), ncol=3)+
  geom_line()+
  labs(title = "Observed genera through time",
       x = "Year of observation",
       y = "number of individuals")+
  theme_bw()+
  theme(text=element_text(size = 16))
```

## Arranging and exporting plots
```{r, fig.width=10,fig.height=7}
library(gridExtra)

spp_weight_boxplot <- ggplot(data = surveys_complete, 
                             aes(x = species_id, y = weight)) +
  geom_boxplot() +
  labs(x = "Species", 
       y = expression(log[10](Weight))) +
  scale_y_log10() + 
  labs()

spp_count_plot <- ggplot(data = yearly_counts, 
                         aes(x = year, y = n, color = genus)) +
  geom_line() + 
  labs(x = "Year", y = "Abundance")

grid.arrange(spp_weight_boxplot, spp_count_plot, ncol = 2, widths = c(4, 6))
```

