---
title: "Lab7"
author: "Alice Zheng"
date: "10/16/2020"
output: html_document
---

### Loading Biocstyle and airway 
```{r}
library("BiocStyle")
library("knitr")
library("rmarkdown")
library("airway")
```

###List files
```{r}
dir <- system.file("extdata", package="airway", mustWork=TRUE)
list.files(dir)
list.files(file.path(dir, "quants"))
```

### Sample Information
```{r}
csvfile <- file.path(dir, "sample_table.csv")
coldata <- read.csv(csvfile, row.names=1, stringsAsFactors=FALSE)
coldata
```

```{r}
coldata <- coldata[1:2,]
coldata$names <- coldata$Run
coldata$files <- file.path(dir, "quants", coldata$names, "quant.sf.gz")
file.exists(coldata$files)
```

```{r}
library("tximeta")
se <- tximeta(coldata)
dim(se)
head(rownames(se))
gse <- summarizeToGene(se)
```


### Looking at gse details
```{r}
dim(gse)
head(rownames(gse))
```

### Creating a summarized experiment object
```{r}
par(mar=c(0,0,0,0))
plot(1,1,xlim=c(0,100),ylim=c(0,100),bty="n",
     type="n",xlab="",ylab="",xaxt="n",yaxt="n")
polygon(c(45,90,90,45),c(5,5,70,70),col="pink",border=NA)
polygon(c(45,90,90,45),c(68,68,70,70),col="pink3",border=NA)
text(67.5,40,"assay(s)")
text(67.5,35,'e.g. "counts", ...')
polygon(c(10,40,40,10),c(5,5,70,70),col="skyblue",border=NA)
polygon(c(10,40,40,10),c(68,68,70,70),col="skyblue3",border=NA)
text(25,40,"rowRanges")
polygon(c(45,90,90,45),c(75,75,95,95),col="palegreen",border=NA)
polygon(c(45,47,47,45),c(75,75,95,95),col="palegreen3",border=NA)
text(67.5,85,"colData")
```

### Load the full gse data
```{r}
data(gse)
gse
```

```{r}
assayNames(gse)
head(assay(gse), 3)
colSums(assay(gse))
```

### Looking at additional gse dataset details
```{r}
rowRanges(gse)
seqinfo(rowRanges(gse))
colData(gse)
```

### gse variables
```{r}
gse$donor
gse$condition
gse$cell <- gse$donor
gse$dex <- gse$condition
```

### renaming gse levels
```{r}
levels(gse$dex)
levels(gse$dex) <- c("untrt", "trt")
```

### loading magrittr and counting reads
```{r}
library("magrittr")
gse$dex %<>% relevel("untrt")
gse$dex
round( colSums(assay(gse)) / 1e6, 1 )
```

### Counts and average transcript lengths
```{r}
library("DESeq2")
dds <- DESeqDataSet(gse, design = ~ cell + dex)
countdata <- round(assays(gse)[["counts"]])
head(countdata, 3)
```

### Converting the counts to integer
```{r}
coldata <- colData(gse)
ddsMat <- DESeqDataSetFromMatrix(countData = countdata,
                                 colData = coldata,
                                 design = ~ cell + dex)
```

### Removing low sample counts
```{r}
nrow(dds)
keep <- rowSums(counts(dds)) > 1
dds <- dds[keep,]
nrow(dds)
keep <- rowSums(counts(dds) >= 10) >= 3
```

### Graphing the variance stabilizing transformation and the rlog
```{r}
lambda <- 10^seq(from = -1, to = 2, length = 1000)
cts <- matrix(rpois(1000*100, lambda), ncol = 100)
library("vsn")
```


```{r}
meanSdPlot(cts, ranks = FALSE)
log.cts.one <- log2(cts + 1)
meanSdPlot(log.cts.one, ranks = FALSE)
```

### Looking at coldata to decide which transformation to choose
```{r}
vsd <- vst(dds, blind = FALSE)
head(assay(vsd), 3)
colData(vsd)
rld <- rlog(dds, blind = FALSE)
head(assay(rld), 3)
```

```{r}
library("dplyr")
library("ggplot2")
```


### Looking at sample distributions
```{r}
sampleDists <- dist(t(assay(vsd)))
sampleDists  
```


### Assess overall similarities using sample distances
```{r}
library("pheatmap")
library("RColorBrewer")
sampleDistMatrix <- as.matrix( sampleDists )
rownames(sampleDistMatrix) <- paste( vsd$dex, vsd$cell, sep = " - " )
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists,
         col = colors)
```

### PCA plot using VST data
```{r}
library("PoiClaClu")
poisd <- PoissonDistance(t(counts(dds)))
samplePoisDistMatrix <- as.matrix( poisd$dd )
rownames(samplePoisDistMatrix) <- paste( dds$dex, dds$cell, sep=" - " )
colnames(samplePoisDistMatrix) <- NULL
pheatmap(samplePoisDistMatrix,
         clustering_distance_rows = poisd$dd,
         clustering_distance_cols = poisd$dd,
         col = colors)
plotPCA(vsd, intgroup = c("dex", "cell"))
```

### PCA plot with custom ggplot2 code
```{r}
pcaData <- plotPCA(vsd, intgroup = c( "dex", "cell"), returnData = TRUE)
pcaData
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = dex, shape = cell)) +
  geom_point(size =3) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed() +
  ggtitle("PCA with VST data")
```

###PCA plot using Generalized PCA
```{r}
library("glmpca")
gpca <- glmpca(counts(dds), L=2)
gpca.dat <- gpca$factors
gpca.dat$dex <- dds$dex
gpca.dat$cell <- dds$cell
ggplot(gpca.dat, aes(x = dim1, y = dim2, color = dex, shape = cell)) +
  geom_point(size =3) + coord_fixed() + ggtitle("glmpca - Generalized PCA")
```



### Building a results table

```{r}
dds <- DESeq(dds)
res <- results(dds)
res
res <- results(dds, contrast=c("dex","trt","untrt"))
mcols(res, use.names = TRUE)
```


```{r}
summary(res)
res.05 <- results(dds, alpha = 0.05)
table(res.05$padj < 0.05)
resLFC1 <- results(dds, lfcThreshold=1)
table(resLFC1$padj < 0.1)
```

### Multiple testings
```{r}
results(dds, contrast = c("cell", "N061011", "N61311"))
sum(res$pvalue < 0.05, na.rm=TRUE)
sum(!is.na(res$pvalue))
sum(res$padj < 0.1, na.rm=TRUE)
resSig <- subset(res, padj < 0.1)
head(resSig[ order(resSig$log2FoldChange), ])
head(resSig[ order(resSig$log2FoldChange, decreasing = TRUE), ])
```


### Plotting results with a counts plot
```{r}
topGene <- rownames(res)[which.min(res$padj)]
plotCounts(dds, gene = topGene, intgroup=c("dex"))
library("ggbeeswarm")
geneCounts <- plotCounts(dds, gene = topGene, intgroup = c("dex","cell"),
                         returnData = TRUE)
```


```{r}
ggplot(geneCounts, aes(x = dex, y = count, color = cell)) +
  scale_y_log10() +  geom_beeswarm(cex = 3)
```

```{r}
ggplot(geneCounts, aes(x = dex, y = count, color = cell, group = cell)) +
  scale_y_log10() + geom_point(size = 3) + geom_line()
```

### MA-plot
```{r}
library("apeglm")
resultsNames(dds)
res <- lfcShrink(dds, coef="dex_trt_vs_untrt", type="apeglm")
plotMA(res, ylim = c(-5, 5))
```


### MA-plot of changes induced by treatment
```{r}
res.noshr <- results(dds, name="dex_trt_vs_untrt")
plotMA(res.noshr, ylim = c(-5, 5))
```

```{r}
plotMA(res, ylim = c(-5,5))
topGene <- rownames(res)[which.min(res$padj)]
with(res[topGene, ], {
  points(baseMean, log2FoldChange, col="dodgerblue", cex=2, lwd=2)
  text(baseMean, log2FoldChange, topGene, pos=2, col="dodgerblue")
})
```


### In histogram form
```{r}
hist(res$pvalue[res$baseMean > 1], breaks = 0:20/20,
     col = "grey50", border = "white")
```

### Gene clustering heatmap of transformed values
```{r}
library("genefilter")
library("pheatmap")
topVarGenes <- head(order(rowVars(assay(vsd)), decreasing = TRUE), 20)

mat  <- assay(vsd)[ topVarGenes, ]
mat  <- mat - rowMeans(mat)
anno <- as.data.frame(colData(vsd)[, c("cell","dex")])
pheatmap(mat, annotation_col = anno)
```


### Independent filtering
```{r}
qs <- c(0, quantile(resLFC1$baseMean[resLFC1$baseMean > 0], 0:6/6))
bins <- cut(resLFC1$baseMean, qs)
levels(bins) <- paste0("~", round(signif((qs[-1] + qs[-length(qs)])/2, 2)))
fractionSig <- tapply(resLFC1$pvalue, bins, function(p)
  mean(p < .05, na.rm = TRUE))
barplot(fractionSig, xlab = "mean normalized count",
        ylab = "fraction of small p values")
```

### Annotating and exporting results
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")

columns(org.Hs.eg.db)

ens.str <- substr(rownames(res), 1, 15)
res$symbol <- mapIds(org.Hs.eg.db,
                     keys=ens.str,
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
res$entrez <- mapIds(org.Hs.eg.db,
                     keys=ens.str,
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")
resOrdered <- res[order(res$pvalue),]
head(resOrdered)
```


### Using SVA with DESeq2
```{r}
library("sva")

dat  <- counts(dds, normalized = TRUE)
idx  <- rowMeans(dat) > 1
dat  <- dat[idx, ]
mod  <- model.matrix(~ dex, colData(dds))
mod0 <- model.matrix(~   1, colData(dds))
svseq <- svaseq(dat, mod, mod0, n.sv = 2)
svseq$sv
par(mfrow = c(2, 1), mar = c(3,5,3,1))
for (i in 1:2) {
  stripchart(svseq$sv[, i] ~ dds$cell, vertical = TRUE, main = paste0("SV", i))
  abline(h = 0)
}
```

### Using RUV with DESeq2
```{r}
library("RUVSeq")

set <- newSeqExpressionSet(counts(dds))
idx  <- rowSums(counts(set) > 5) >= 2
set  <- set[idx, ]
set <- betweenLaneNormalization(set, which="upper")
not.sig <- rownames(res)[which(res$pvalue > .1)]
empirical <- rownames(set)[ rownames(set) %in% not.sig ]
set <- RUVg(set, empirical, k=2)
pData(set)

par(mfrow = c(2, 1), mar = c(3,5,3,1))
for (i in 1:2) {
  stripchart(pData(set)[, i] ~ dds$cell, vertical = TRUE, main = paste0("W", i))
  abline(h = 0)
}
```


### Time course experiments
```{r}
library("fission")
data("fission")
ddsTC <- DESeqDataSet(fission, ~ strain + minute + strain:minute)
ddsTC <- DESeq(ddsTC, test="LRT", reduced = ~ strain + minute)
resTC <- results(ddsTC)
resTC$symbol <- mcols(ddsTC)$symbol
head(resTC[order(resTC$padj),], 4)

fiss <- plotCounts(ddsTC, which.min(resTC$padj), 
                   intgroup = c("minute","strain"), returnData = TRUE)
fiss$minute <- as.numeric(as.character(fiss$minute))
ggplot(fiss,
       aes(x = minute, y = count, color = strain, group = strain)) + 
  geom_point() + stat_summary(fun.y=mean, geom="line") +
  scale_y_log10()
```

### Fission heatmap
```{r}
resultsNames(ddsTC)
res30 <- results(ddsTC, name="strainmut.minute30", test="Wald")
res30[which.min(resTC$padj),]

betas <- coef(ddsTC)
colnames(betas)

topGenes <- head(order(resTC$padj),20)
mat <- betas[topGenes, -c(1,2)]
thr <- 3 
mat[mat < -thr] <- -thr
mat[mat > thr] <- thr
pheatmap(mat, breaks=seq(from=-thr, to=thr, length=101),
         cluster_col=FALSE)
```


 
