---
title: "Lab 9"
author: "Alice Zheng"
date: "10/26/2020"
output: html_document
---
```{r}
library(stringr)
library(tidyverse)
```

### 14.3.1.1
these strings don't match a \\\ because you need 4 \\\'s in order to match one backslash

```{r}
x <- "a\\b"
str_view(x, "\\\n")
```


### 14.3.2.1
```{r, results="hide"}
str_view(x, "^\\$\\^\\$$")
```


### 14.3.3.1
```{r, results="hide"}
str_view(x, "^[aeiou]")
str_view(x, "^[^aeiou]")
str_view(x, "[^e]ed$")
str_view(x, "[ing|ise]$")
```



### 14.3.4.1
? {0,1}
+ {1,}
* {0,}

### 14.3.5.1
a character repeated 2 more times
a pair of characters with the reverse after it
a a pair of characters repeated once
a character followed by any character, the character again, any character, and the character once more
three characters, 0 or more characters of any kind, the three characters in reverse order

### 14.4.1.1

```{r, results="hide"}
str_detect(words, "^x|x$") 
start <- str_detect(words, "^x")
end <- str_detect(words, "x$")
words[start|end]
```

```{r, results="hide"}
str_detect(words, "^[aeiou]|[^aeiou]$") 
start <- str_detect(words, "^[aeiou]")
end <- str_detect(words, "[^aeiou]$")
words[start|end]
```


```{r}
a <- str_detect(words, "a") 
ae <- str_detect(a, "e") 
aei <- str_detect(ae, "i") 
aeio <- str_detect(aei, "o") 
aeiou <- str_detect(aeio, "u") 
head(aeiou)
```

### 14.4.2.1
```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```

### 14.4.3.1
```{r}
num <- "(one|two|three|four|five|six|seven|eight|nine|ten) ([^ ]+)"
has_num <- sentences %>%
  str_subset(num) %>%
  head(10)
has_num %>% 
  str_extract(num)
```

```{r}
has_num %>% 
  str_match(num)
```

### 14.4.4.1
```{r}
s <- "\\\\\\"
s1 <- str_replace_all(s, "\\\\", "/")
s1
```

### 14.4.5.1
```{r}
s <- "apples, pears, and bananas"
str_split(s, boundary("word"))[[1]]
```




